---
title: "TARI Bean Program Workshop: GxE Analysis in R"
draft: true
description: |
  Day x of the TARI Bean Program Training hosted by CIAT Arusha.
author:
  - name: Ellena Girma
    url: "https://rforplantbreeders.netlify.app/about.html"
date: "`r Sys.Date()`"
output: distill::distill_article
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
library(xaringan)
library(xaringanExtra)
library(xaringanthemer)
library(xaringanBuilder)
library(pagedown)
```


### Welcome

My name is Ellena Girma and I am a data analyst intern at CIAT Arusha. I will be leading today's workshop on GxE Interaction and Stability Analysis. 

The goal of this workshop is to give TARI Bean Project scientists and technicians an overview of how to use R to run GxE Interaction and Stability Analysis.

```{r share-again, echo=FALSE}
xaringanExtra::use_share_again()
```

```{r embed-xaringan, echo=FALSE, fig.cap="[Slides](gxe-analysis-June2022)"}
xaringanExtra::embed_xaringan(url = "https://ellenagirma-workshops.netlify.app/gxe-june2022/", ratio = "8:5")
```


### Materials 

#### R

You will need to install R as well as the following packages:

```{r installing-packages}

#install.packages(c("tidyverse", "lme4", "metan"))

```

#### Data

Feel free to use your own data if you have data available. If you need data to practice on, I suggest using data from <a href = "https://www.sciencedirect.com/science/article/pii/S2405844021000785">this</a> paper. ^[''Philipo M, Ndakidemi PA, Mbega ER. Environmentally stable common bean genotypes for production in different agro-ecological zones of Tanzania. Heliyon. 2021 Jan 19;7(1):e05973. doi: 10.1016/j.heliyon.2021.e05973''
] The data contains information about the seed yield, 100 seed weight, number of seeds/pod, number of pods/plant and days to 75% flowering for 99 common bean variants grown in 3 locations. The data is listed under Appendix A. Supplementary data. 

