---
title: "Bean Program Workshop: GxE Analysis in R"
draft: false
description: |
  GxE Analysis in R Workshop hosted by CIAT Arusha. 
  Updated `r Sys.Date()`
author:
  - name: Ellena Girma
    url: "https://rforplantbreeders.netlify.app/about.html"
date: July 10, 2022
output: distill::distill_article
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
library(xaringan)
library(xaringanExtra)
library(xaringanthemer)
library(renderthis)
library(pagedown)
```


### Welcome

My name is Ellena Girma and I am a breeding data systems and analytics specialist at the Alliance Bioversity International - CIAT.

The goal of this workshop is to give CIAT Bean Program scientists and technicians an overview of how to use the metan pacakge in R to run GxE Interaction and Stability Analysis.

```{r share-again}
xaringanExtra::use_share_again()

```

```{r embed-xaringan, echo=FALSE, fig.cap="[Slides](slides/index.html)", include=FALSE}
xaringanExtra::embed_xaringan(url = "https://ellenagirma-workshops.netlify.app/gxe-june2022/index.html", ratio = "16:9")
```


### Materials 

Below are the materials you need to follow the workshop:

#### <a href = "https://ellenagirma-workshops.netlify.app/gxe-june2022/">Slides</a>

#### Installing R & RStudio

You will need to install R/RStudio* as well as the following packages:

* tidyverse
* metan

Check out my <a href = "https://rforplantbreeders.netlify.app/resources.html">resources</a> page for video instructions on how to download R and RStudio. 

#### Project-Oriented Workflow

It's best to organize your R code using R Projects. This way, all the materials related to an analysis/project will be under one parent folder but will be separated by content type (ex: data, scripts, outputs, reports, notes). All folders in this parent folder are easily accessed by the RProject. 

Some resources on project oriented workflows in R:

  * <a href = "https://rstats.wtf/project-oriented-workflow.html">What They Forgot to Teach You About R: "Project-oriented workflow" chapter by Jenny Bryan and Jim Hester</a>
  * <a href = "https://r4ds.had.co.nz/workflow-projects.html">R for data science: "Workflow: projects" chapter by Hadley Wickham</a>
  * <a href = "https://www.tidyverse.org/blog/2017/12/workflow-vs-script/">"Project-oriented workflow" blog post by Jenny Bryan</a>

#### Data

Feel free to use your own data if you have data available. 

If you need data to practice on, I suggest using data from <a href = "https://www.sciencedirect.com/science/article/pii/S2405844021000785">this</a> paper. ^[''Philipo M, Ndakidemi PA, Mbega ER. Environmentally stable common bean genotypes for production in different agro-ecological zones of Tanzania. Heliyon. 2021 Jan 19;7(1):e05973. doi: 10.1016/j.heliyon.2021.e05973''
] The data contains information about the seed yield, 100 seed weight, number of seeds/pod, number of pods/plant and days to 75% flowering for 99 common bean variants grown in 3 locations. The data is listed under Appendix A. Supplementary data. 

